# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

Load knitr and ensure code chunks "echo"

```{r stage}
library(knitr)
opts_chunk$set(echo=TRUE)
```

Get the CSV file. Transform to data table. Convert date from "factor" to "Date"

```{r load}
library(data.table)
AT <- fread("activity/activity.csv", sep=",")
AT[,date:=as.Date(date,format="%Y-%m-%d")]
```

## What is mean total number of steps taken per day?

Plot a histogram of totals steps per day:

```{r histsteps}
s <- tapply(AT[,steps],AT[,date],sum,na.rm=T)
hist(s,main="Histogram of steps per day",xlab="Steps per day")
```

Give the mean and median steps per day:

```{r averages}
mean(s)
median(s)
```


## What is the average daily activity pattern?

```{r TypicalDay}
t <- tapply(AT[,steps],AT[,interval],mean,na.rm=T)
plot(names(t),t,type="l",main = "Average Daily Activity Pattern",xlab="5-minute interval",ylab="steps")
```

Which interval, on average, contains the highest number of steps?

```{r TypicalInterval}
names(t[t==max(t)])
```

## Imputing missing values

How many values are missing? Impute the missing values and create a dataset including the imputed values where the missing values were.

```{r missing}
length(AT[is.na(AT$steps),][[1]])
AT[,intmean:=round(mean(steps,na.rm=T)),by=interval]
AT[is.na(steps),steps:=intmean]
```

Plot a histogram and show mean and median values for the dataset with the imputed values. Compare to the values from the original data:

```{r althiststeps}
s <- tapply(AT[,steps],AT[,date],sum,na.rm=T)
hist(s,main="Histogram of steps per day including imputed values",xlab="Steps per day")
mean(s)
median(s)
```

## Are there differences in activity patterns between weekdays and weekends?

```{r wkdy}
AT[,wkdy:=weekdays(date)]
AT[wkdy %in% c("Saturday","Sunday"),daytype:="Weekend"]
AT[wkdy %in% c("Monday","Tuesday","Wednesday","Thursday","Friday"),daytype:="Weekday"]
ATwkday <- AT[daytype == "Weekday",]
ATwkend <- AT[daytype == "Weekend",]
wdayavgsteps <- tapply(ATwkday[,steps],ATwkday[,interval],mean)
wendavgsteps <- tapply(ATwkend[,steps],ATwkend[,interval],mean)
xlabels <- names(wendavgsteps)
par(mfrow=c(2,1))
plot(xlabels,wendavgsteps,type="l",main="Weekend",xlab="",ylab="")
plot(xlabels,wdayavgsteps,type="l",main="Weekday",xlab="Interval",ylab="Avg Steps")
dev.off()
```


